A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa

### 2019-10-16 literature_review.Rmd
began section on particle dispersion modelling. inserted complicated equation. with another one to follow

---
title: A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa
author: "Ross Coppin"
output:
  pdf_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
bibliography: references.bib
---

@@

An object will only float if the object density ($\rho$) is less than or equal to the density of the liquid in which the object is submerged, i.e. $\rho_{object} < \rho_{liquid}$. The bouyancy force (*BF*) is the force of the displaced liquid pushing the object in an upwards direction which is equal to the weight of the displaced liquid, $BF = W_{displaced-liquid}$. The weight of the displaced liquid can be calculated by multiplying the mass of the liquid by the gravitational constant $g$ and so the weight of the displaced liquid can be calculated by: $$W_{displaced-liquid} = mg$$ The equation is expanded upon by including the calculation for mass, $m = \rho V$, and so now the weight of the displaced liquid can be calculated as: $$mg = \rho_{liquid} V g$$ where $\rho$ represents the density of the liquid and $V$ represents the portion of the object submerged. Since the object is floating, the BF can also be expressed by the weight of the object and as mentioned previously the weight of the displaced liquid can also be expressed by $mg$, which in turn can be written as $W_{displaced-liquid} = W_{object}$. Now through substituting the calculations for weight, volume and the dimensions of the object, the equation can be simplied to: $$x= h\displaystyle\frac{\rho_o}{\rho_l}$$ where $h$ is the overall height of the object, $\rho_o$ represents the density of the object and $\rho_l$ represents the density of the liquid. It is important to note that this equation is only applicable to objects that are floating i.e. $\rho_{object} < \rho_{liquid}$. The portion of a floating object above the water surface can also be calcuated using a similar approach, and instead of calculating the submerged portion ($x$) the calculation will now solve for the fraction above the surface which is represented by $y$ (see figure \ref{fig:BF_1}). The same equation for calculating density ($\rho = \displaystyle\frac{m}{V}$) will also be used in solving for $y$. As mentioned previously, for floating objects the $BF = m_{object}g$ and the BF can also be defined as the weight of the displaced liquid ($BF = W_{displaced-liquid}$) and therefore the weight of the displaced liquid is equal to the weight of the object ($W_{displaced-liquid} = W_{object}$). On this basis equation ?? can be written as $$m_{liquid}g = m_{object}g$$
Substituting the equation for density into equation ?? the equation is now: $$\rho_{o}V_{o}g = \rho_{l}V_{l}g$$. Important to note is that $V_{o}$ represents the volume of displaced liquid i.e not the entire object volume (see figure \ref{fig:BF_1}). The constant $g$ is removed through cancellation and now the equation can be written as: $$\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$$ We also now that the volume of the object submerged is equal to the volume of the displaced liquid and the fraction $\displaystyle\frac{V_l}{V_o}$ represents the portion of the object submerged. Therefore to calculate the portion above the surface this must be subtracted by 1, therefore: $$1-\displaystyle\frac{V_l}{V_o} = y$$
and since $\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$ the equation can now be solved via density of the liquid and the object: $$y = 1 - \displaystyle\frac{\rho_o}{\rho_l}$$ Solving the equation algebraically it can now be written as:$$ y = \displaystyle\frac{\rho_l - \rho_o}{\rho_l}$$

@@

The Pol3DD model is able to track virtual particles to simulate water borne dispersion of material. Examples include neutrally buoyant anthropogenic material, larvae, oil spills, outfall discharges and sediment transport [@lebreton2012]. The model uses a second-order accurate advection scheme [@black1990] described as follows:

>>>>

---
title: A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa
author: "Ross Coppin"
output:
  word_document:
    fig_caption: yes
    toc: yes
  pdf_document:
    toc: yes
bibliography: references.bib
---

@@

An object will only float if the object density ($\rho$) is less than or equal to the density of the liquid in which the object is submerged, i.e. $\rho_{object} < \rho_{liquid}$. The bouyancy force (*BF*) is the force of the displaced liquid pushing the object in an upwards direction which is equal to the weight of the displaced liquid and is expressed as: $$BF = W_{displaced-liquid}$$ The weight of the displaced liquid can be calculated by multiplying the mass of the liquid by the gravitational constant $g$ and so the weight of the displaced liquid can be calculated by: $$W_{displaced-liquid} = mg$$ The equation is expanded upon by including the calculation for mass, $m = \rho V$, and so now the weight of the displaced liquid can be calculated as: $$mg = \rho_{liquid} V g$$ where $\rho$ represents the density of the liquid and $V$ represents the portion of the object submerged. Since the object is floating, the BF can also be expressed by the weight of the object and as mentioned previously the weight of the displaced liquid can also be expressed by $mg$, which in turn can be written as $W_{displaced-liquid} = W_{object}$. Now through substituting the calculations for weight, volume and the dimensions of the object, the equation can be simplied to: $$x= h\displaystyle\frac{\rho_o}{\rho_l}$$ where $h$ is the overall height of the object, $\rho_o$ represents the density of the object and $\rho_l$ represents the density of the liquid. It is important to note that this equation is only applicable to objects that are floating i.e. $\rho_{object} < \rho_{liquid}$. The portion of a floating object above the water surface can also be calcuated using a similar approach, and instead of calculating the submerged portion ($x$) the calculation will now solve for the fraction above the surface which is represented by $y$ (see figure \ref{fig:BF_1}). The same equation for calculating density ($\rho = \displaystyle\frac{m}{V}$) will also be used in solving for $y$. As mentioned previously, for floating objects the $BF = m_{object}g$ and the BF can also be defined as the weight of the displaced liquid ($BF = W_{displaced-liquid}$) and therefore the weight of the displaced liquid is equal to the weight of the object ($W_{displaced-liquid} = W_{object}$). On this basis equation ?? can be written as $$m_{liquid}g = m_{object}g$$
Substituting the equation for density into equation ?? the equation is now: $$\rho_{o}V_{o}g = \rho_{l}V_{l}g$$ Important to note is that $V_{o}$ represents the volume of displaced liquid i.e not the entire object volume (see figure \ref{fig:BF_1}). The constant $g$ is removed through cancellation and now the equation can be written as: $$\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$$ We also now that the volume of the object submerged is equal to the volume of the displaced liquid and the fraction $\displaystyle\frac{V_l}{V_o}$ represents the portion of the object submerged. Therefore to calculate the portion above the surface this must be subtracted by 1, therefore: $$1-\displaystyle\frac{V_l}{V_o} = y$$
and since $\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$ the equation can now be solved via density of the liquid and the object: $$y = 1 - \displaystyle\frac{\rho_o}{\rho_l}$$ Solving the equation algebraically it can now be written as:$$ y = \displaystyle\frac{\rho_l - \rho_o}{\rho_l}$$

@@

The Pol3DD model is able to track virtual particles to simulate water borne dispersion of material. Examples include neutrally buoyant anthropogenic material, larvae, oil spills, outfall discharges and sediment transport [@lebreton2012]. The model uses a second-order accurate advection scheme [@black1990] described as follows: $$\delta\chi=\frac{({u'+ \frac{(u_y\upsilon'-\upsilon_yu')\delta t}{2} )\delta t}}{(1-\frac{u_x\delta t}{2})(1-\frac{\upsilon_y\delta t}{2})-\frac{u_y\upsilon_\chi\delta t^2}{4}}$$
